---
title: "Flyson/Flynn/Golde/Kowalsky Secret Santa 2022"
author: "Erin Larson"
date: "2022-11-22"
output: rmarkdown::html_document
---
Note: this code was modified from: https://github.com/tjmahr/tjmahr.github.io/blob/master/_R/2017-11-28-secret-santa-graph-traversal.Rmd 

# R set up
Bringing in relevant packages.
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(DiagrammeR)
library(magrittr)
library(dplyr, warn.conflicts = FALSE)
```


# Nibling set up
Creating table of names and family relations.
```{r}
niblings <- tibble::tribble(
      ~Family, ~ Name, ~ Number,
    "Flyson", "Tom",        1,
      "Flyson",  "Erin",        2, 
 "Flynn", "Ryan",        3, 
   "Flynn",   "Alli",        4, 
   "Golwalsky", "Gae",        5,
 "Golwalsky", "Tim", 6
)
```


# Graph Rendering
Function to show gift giving network.
```{r}
render_graph <- function(graph, height = 300, ...) {
  g <- DiagrammeR::render_graph(graph, height = height, ...)
  
  svg_dim <- sprintf("svg width=\"%s\" height=\"%s\"", "100%", height)
  
  DiagrammeRsvg::export_svg(g) %>% 
    stringr::str_replace("^<[?]xml.+\n", "") %>%
    stringr::str_replace("^<[!]DOCTYPE.+\n.+\n", "") %>%
    stringr::str_replace("svg width=\"\\d+pt\" height=\"\\d+pt\"", 
                         svg_dim) %>%
    htmltools::HTML() %>%
    knitr::asis_output()
}
if (interactive()) render_graph <- DiagrammeR::render_graph
```

# Example
Quick example showing what it would look like if Erin drew Gae from a "hat".
```{r}
nodes <- create_node_df(
  n = nrow(niblings),
  type = niblings$Name,
  label = niblings$Name
)
erin_drew_gae <- create_edge_df(
  from = 2, 
  to = 5, 
  rel = "gives-to",
  color = "#FF4136",
  penwidth = 1
)
create_graph(nodes, erin_drew_gae) %>% 
  render_graph()
```

# All possible connections
Creating a graph showing all possible gift-giving connections.
```{r}
combn(niblings$Name, 2)
# All the edge-manipulating functions in this post take an optional take `...`
# argument for setting the style of edges.
create_all_giving_edges <- function(xs, ...) {
  aes_options <- quos(...)
  pairs <- combn(seq_along(xs), 2)
  # Each column from `combn()` is a pair. We make an edge moving down the column
  # and another edge up the column by having each row as a `from` index and a
  # `to` index.
  from <- c(pairs[1, ], pairs[2, ])
  to <- c(pairs[2, ], pairs[1, ])
  create_edge_df(from = from, to = to) %>% 
    mutate(!!! aes_options) %>% 
    as_tibble()
}
all_possible_edges <- create_all_giving_edges(
  niblings$Name, 
  rel = "potential-gift", 
  penwidth = .5,
  color = "#CCCCCC90"
)
create_graph(nodes, all_possible_edges) %>% 
  render_graph()
```
# Gift giving dataframe
Creating a table with all possible combinations of gift-giving.
```{r}
nibling_edges <- niblings$Name %>% 
  create_all_giving_edges(
    rel = "potential-gift", 
    penwidth = .5,
    color = "#CCCCCC90"
  ) %>% 
  left_join(niblings, by = c("from" = "Number")) %>% 
  rename(from_fam = Family) %>% 
  select(-Name) %>% 
  left_join(niblings, by = c("to" = "Number")) %>% 
  rename(to_fam = Family) %>% 
  select(-Name) %>% 
  select(id, from, to, rel, from_fam, to_fam, everything())
nibling_edges
```
# Partner graph
Creating a graph showing partner pairs.

```{r}
overwrite_edges <- function(old_df, new_df) {
  old_df %>%
    anti_join(new_df, by = c("from", "to")) %>%
    bind_rows(new_df)
}

partner_edges <- nibling_edges %>% 
  filter(from_fam == to_fam) %>% 
  mutate(
    rel = "partner",
    color = "#3D9970", 
    penwidth = 1)
# Update edges that represent partner pairs
  
nibling_nodes <- create_node_df(
  n = nrow(niblings),
  type = niblings$Name,
  label = niblings$Name)
nibling_edges %>% 
  overwrite_edges(partner_edges) %>% 
  create_graph(nibling_nodes, .) %>% 
  render_graph(height = 400)
```
# Some useful functions
Functions to forbid partner pairs from gifting to each other and the drawing a name from a hat function.
```{r}
has_free_edge <- function(edge_df) {
  edges_left <- edge_df %>% filter(rel != "gives-to") %>% nrow()
  edges_left != 0
}

draw_secret_santa_edge <- function(edge_df, ...) {
  aes_options <- quos(...)
  
  edge_df %>%
    filter(rel != "gives-to") %>%
    sample_n(1) %>%
    mutate(!!! aes_options)
}
find_illegal_edges <- function(edge_df, edge, ...) {
  aes_options <- quos(...)
  
  outgoing <- edge_df %>%
    filter(from %in% edge$from)
  incoming <- edge_df %>%
    filter(to %in% edge$to)
  # The one edge that is not illegal is in both 
  # the incoming and outgoing sets
  to_keep <- dplyr::intersect(outgoing, incoming)
  outgoing %>% 
    bind_rows(incoming) %>% 
    anti_join(to_keep, by = c("from", "to")) %>% 
    mutate(!!! aes_options)
}

has_valid_gift_edges <- function(edge_df, indices) {
  indices <- sort(unique(indices))
  pairs <- edge_df %>% filter(rel == "gives-to")
  no_self_loop <- !any(pairs$from == pairs$to)
  exhaustive_from <- isTRUE(all.equal(sort(pairs$from), indices))
  exhaustive_to <- isTRUE(all.equal(sort(pairs$to), indices))
  all(no_self_loop, exhaustive_from, exhaustive_to)
}

draw_edges_from_hat <- function(edge_df, ...) {
  aes_options <- quos(...)
  raw_edge_df <- edge_df
  indices <- unique(c(raw_edge_df$from, raw_edge_df$to))
  
  while (has_free_edge(edge_df)) {
    pick <- edge_df %>% 
      draw_secret_santa_edge(!!! aes_options) %>% 
      mutate(rel = "gives-to")
    
    illegal_edges <- edge_df %>%
      find_illegal_edges(pick)
    edge_df <- edge_df %>%
      overwrite_edges(pick) %>% 
      anti_join(illegal_edges, by = "id") 
  }
  if (!has_valid_gift_edges(edge_df, indices)) {
    warning(call. = FALSE, "Invalid drawing. Trying again.")
    edge_df <- Recall(raw_edge_df, !!! aes_options)
  }
  
  edge_df
}

```

# The draw!
Drawing names from a hat, forbidding partners to give/receive gifts from each other.
```{r}
nibling_edges <- nibling_edges %>% 
  overwrite_edges(partner_edges)
  
nibling_nodes <- create_node_df(
  n = nrow(niblings),
  type = niblings$Name,
  label = niblings$Name)
nibling_edges %>% 
  overwrite_edges(partner_edges) %>% 
  create_graph(nibling_nodes, .) %>% 
  render_graph(height = 400)


nibling_edges %>% 
  filter(rel != "partner") %>% 
  draw_edges_from_hat(color = "#FF4136") %>% 
  create_graph(nibling_nodes, .) %>% 
  render_graph(height = 500)
```